Draft-Theory

### 2024-11-16 descriptivestry.Rmd
darkmodeleggo

>>>>

<!DOCTYPE html>
<html lang="en">
<head>
    <meta charset="UTF-8">
    <meta name="viewport" content="width=device-width, initial-scale=1.0">
    <title>My Lab Journal</title>
    <link rel="stylesheet" href="styles.css">  <!-- Link to the CSS file I created for dark mode code chunks -->

</head>
<body>
    <header>
        <h1>Welcome to My Lab Journal</h1>
        <p>This is where I record my research and experiments.</p>
    </header>

    <main>
        <section>
            <h2>Experiment 1</h2>
            <p>Details of the first experiment go here.</p>
        </section>
        <section>
            <h2>Experiment 2</h2>
            <p>Details of the second experiment go here.</p>
        </section>
    </main>

    <footer>
        <p>Created by Me, 2024</p>
    </footer>
</body>
</html>

### 2024-11-16 draftdata.Rmd
darkmodeleggo

---
title: "draft-data"
author: "Verooo"
date: "2024-11-01"
output: html_document
bibliography: references.bib
---

@@

For simplicity's sake the concept of gender is not only inferred by the scholars name, but also assumed to be binary and indistinguishable from biological sex. I am aware that names do not equate to someone's gender identity and that gender identities encompass more than men and women. Further I am also aware that the terms men and women refer to gender identities and the terms male and female usually refer to biological sex. In this study these words have been conflated and it should be noted that this is not representative of the current scientific understanding of biological sex and gender, and that the way of obtaining this variable disregards scholars' self-identification of gender completely. This is done merely due to this project being part of a university assignment and is not something I would do, if I were to research outside of the context of my university curriculum.

@@

# written references (remove in final paper)

>>>>

---
title: "draft-data"
author: "Verooo"
date: "2024-11-01"
output: html_document
bibliography: references.bib
csl: apa.csl
---

@@

For the sake of simplicity, gender in this study was inferred from scholars’ names and treated as binary, conflating gender identity with biological sex. I acknowledge that names do not reliably indicate a person's gender identity and that gender encompasses more than the categories of men and women. Additionally, I recognize that while *men* and *women* typically refer to gender identities, *male* and *female* are often used to denote biological sex. These distinctions were not upheld in this study, which conflates the terms and disregards individuals' self-identified genders. This approach was taken solely due to the constraints of a university assignment and does not reflect my standard research practices, which would respect and account for the complexity of gender and self-identification outside the academic coursework context.

@@

# Ethical concerns

## Webscraping

This study relies on data obtained through web scraping publicly available sources, which raises several ethical concerns. One key issue is that the scholars included in the dataset did not provide informed consent for the use of their data in this research. To address this, several measures were implemented. First, efforts were made to anonymize the data presented in this study to the greatest extent possible. Second, only information directly relevant to the research objectives was collected, minimizing potential intrusions on privacy. Third, the scraped data did not include sensitive or harmful content, thereby reducing the risk of any negative consequences arising from its use.

Despite these precautions, the fundamental issue of using non-consensual data persists. I do not adhere to the view that making data publicly available equates to granting unrestricted permission for its use in research. While the data in this study was already accessible online, the research generates new insights, such as patterns of tie formation between scholars, which were not previously explicit. Although names and other identifying details have been omitted to avoid recognizing individuals within network graphs or other outputs, informed consent for the use of personal data was not obtained. As such, this method of data collection remains ethically contentious at best.

## Operationalization of gender

An additional ethical concern relates to the treatment of gender in this study. The method used to infer gender, based on names and a binary framework, may raise issues regarding inclusivity and respect for individuals’ self-identified gender. This approach disregards the complexity of gender identity and risks misrepresentation. A more detailed explanation of how gender data was defined and its limitations is provided in the data definition section.

# Written References (remove in final paper)

### 2024-11-16 draftintro.Rmd
darkmodeleggo

citation_package: apacite
bibliography: references.bib
biblio-style: apacite
natbiboptions: round
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

Collaboration has become an essential part of modern academia, with scholars increasingly engaging in joint research efforts. Co-authorship allows researchers to combine expertise, share resources, and enhance their scholarly output (@Cheong2009). However, the dynamics of academic collaboration are not free from social patterns, and gender homophily—where individuals prefer to collaborate with others of the same gender—may play a significant role. Understanding how gender affects co-authorship networks, particularly in terms of homophily, can reveal important insights into how academic opportunities and achievements are distributed across genders.

Previous studies have found that male researchers often prefer collaborating with other men, while female researchers tend to collaborate with other women, albeit to a lesser extent (@Horta2022). This raises critical questions about the structure and inclusiveness of academic networks. In fields such as Sociology and Political Science, where collaboration is crucial for scholarly success, gender-based patterns could shape the career trajectories of researchers (@Cheong2009). Furthermore, as Kwiek and Roszka (@Kwiek2021) point out, all-male authorship is far more prevalent than all-female authorship, indicating that male-dominated collaboration networks may reinforce existing gender imbalances in academia.

@@

@@

>>>>

bibliography: references.bib
csl: apa.csl
---

```{=html}
<style>
body {
text-align: left;
font-family: Times;
}

h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
  font-family: Times;
}

</style>
```

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# intro second draft :
# Introduction

## Research Problem and Goal

In this study, I examine the role of gender homophily in shaping co-authorship networks in the Sociology and Political Science Departments at Radboud University. Gender homophily refers to the tendency of individuals to collaborate with others of the same gender. This phenomenon can influence who gains access to collaboration opportunities, thereby shaping academic trajectories and reinforcing existing inequalities. Understanding these dynamics is essential to uncovering how collaboration networks may perpetuate or mitigate gender-based disparities in academia.

### Gender and Collaboration

Previous research has consistently shown that gender plays a significant role in academic collaboration. Studies by Horta (2022) and Kwiek and Roszka (2021) reveal that male researchers are more likely to collaborate with other men, while female researchers exhibit weaker but still present homophilous tendencies. This pattern contributes to the persistence of gender imbalances in co-authorship, where all-male teams remain more common than all-female teams. These trends are particularly concerning in fields such as Sociology and Political Science, where collaboration is vital for career advancement.

The structure of these networks affects not only who collaborates but also how academic recognition and resources are distributed. For example, male-dominated networks may facilitate better access to high-impact collaborations and grant opportunities for men, further perpetuating the gender gap in academic success.

### Prestige and Collaboration

Another crucial factor influencing collaboration is academic prestige, often measured by publication in high-impact journals. Scholars with higher prestige, indicated by metrics such as the h-index or publication in top-tier journals, are generally more sought after as collaborators. Existing research suggests that prestige might interact with gender homophily in significant ways. Bellotti et al. (2022) found that prestigious male scholars often form networks that exclude women, while McDowell, Singell, and Stater (2006) observed that female scholars, despite demonstrating similar academic competence, face greater challenges in accessing high-prestige collaborative networks.

In this context, it is crucial to explore whether the tendency for gender-based collaboration shifts when prestige factors come into play. Specifically, does publishing in prestigious journals mitigate or exacerbate gender homophily in academic networks?

## Research Focus and Questions

This study aims to analyze gender homophily within the co-authorship networks of the Sociology and Political Science Departments at Radboud University, focusing on the role of journal prestige. The research is guided by the following questions:

- **RQ1a:** To what extent does gender homophily occur across all co-publications of the staff in the Sociology and Political Science Departments at Radboud University?

- **RQ1b:** To what extent do the two departments differ in their levels of gender homophily?

These questions focus on descriptive patterns, providing a macro-level view of gender clustering in academic networks.

- **RQ2:** To what extent can gender homophily in collaboration networks be explained by preferences at the dyadic level, controlling for structural network and individual-level covariates?

This question shifts the focus to the micro-level, examining whether individual choices and structural factors drive homophily in co-authorship.

- **RQ3:** How does the prestige of scholars, measured by their publication in high- versus low-impact journals, influence gender homophily in co-authorship networks?

This question explores whether prestige interacts with gender homophily, shedding light on how academic recognition impacts collaborative behaviors. Specifically, it investigates whether male and female scholars with similar prestige levels experience different patterns of collaboration and whether high-prestige contexts reduce or amplify gender clustering.

## Relevance and Contribution

This study contributes to the growing body of literature on gender inequalities in academia by focusing on the intersection of gender and academic prestige in shaping collaboration networks. By applying a social network perspective, it offers insights into how gender-based preferences influence co-authorship patterns and how these dynamics vary between disciplines. Moreover, by exploring the role of journal prestige, this research highlights how structural factors may intersect with social biases to shape academic opportunities, providing a nuanced understanding of the mechanisms underlying gender disparities in academia.

---

@@

## Contextualization
Collaboration has become a cornerstone of modern academic practice, with scholars increasingly working together to leverage diverse expertise, share resources, and enhance their research output. Co-authorship, in particular, is a key indicator of collaboration, often reflecting the dynamics of professional relationships and the distribution of academic opportunities. As Cheong (@Cheong2009) highlights, collaboration enables researchers to combine intellectual and institutional resources, contributing significantly to their scholarly success.

However, collaboration is not free from social patterns and biases. Academic networks often mirror broader societal inequalities, with persistent disparities in opportunities and recognition. One such disparity is gender inequality, which continues to shape interactions, opportunities, and outcomes within academia.

## Research problem & goal

Understanding how gender affects co-authorship networks, particularly in terms of homophily, can reveal important insights into how academic opportunities and achievements are distributed across genders.
## N: Research Problem and Goal
In this study, I examine the role of gender homophily in shaping co-authorship networks in the Sociology and Political Science Departments at Radboud University. Gender homophily refers to the tendency of individuals to collaborate with others of the same gender. This phenomenon can influence who gains access to collaboration opportunities, thereby shaping academic trajectories and reinforcing existing inequalities. Understanding these dynamics is essential to uncovering how collaboration networks may perpetuate or mitigate gender-based disparities in academia.

## Short overview of prior findings:

### gender & collaboration

Previous studies have found that male researchers often prefer collaborating with other men, while female researchers tend to collaborate with other women, albeit to a lesser extent (@Horta2022). This raises critical questions about the structure and inclusiveness of academic networks. In fields such as Sociology and Political Science, where collaboration is crucial for scholarly success, gender-based patterns could shape the career trajectories of researchers (@Cheong2009). Furthermore, as Kwiek and Roszka (@Kwiek2021) point out, all-male authorship is far more prevalent than all-female authorship, indicating that male-dominated collaboration networks may reinforce existing gender imbalances in academia.
### N:Gender and Collaboration

Previous research has consistently shown that gender plays a significant role in academic collaboration. Studies by Horta (2022) and Kwiek and Roszka (2021) reveal that male researchers are more likely to collaborate with other men, while female researchers exhibit weaker but still present homophilous tendencies. This pattern contributes to the persistence of gender imbalances in co-authorship, where all-male teams remain more common than all-female teams. These trends are particularly concerning in fields such as Sociology and Political Science, where collaboration is vital for career advancement.

The structure of these networks affects not only who collaborates but also how academic recognition and resources are distributed. For example, male-dominated networks may facilitate better access to high-impact collaborations and grant opportunities for men, further perpetuating the gender gap in academic success.

### h-index & collaboration

@@

### Reiteration of (societal) relevance of study

@@

# Written references (remove in final paper)

Bornmann, Lutz, Christian Ganser, and Alexander Tekles. 2021. “Simulation of the h Index Use at University Departments Within the Bibliometrics-Based Heuristics Framework: Can the Indicator Be Used to Compare Individual Researchers?” *Journal of Informetrics* 16 (1): 101237. <https://doi.org/10.1016/j.joi.2021.101237>.

@Bornmann2021

Bornmann, Lutz, Christian Ganser, Alexander Tekles, and Loet Leydesdorff. 2020. “Does the h αα-Index Reinforce the Matthew Effect in Science? The Introduction of Agent-Based Simulations into Scientometrics.” *Quantitative Science Studies* 1 (1): 331–46.

@Bornmann2020

### 2024-11-16 drafttheory.Rmd
darkmodeleggo

---
title: "Draft-Theory"
author: "Verooo"
date: "2024-10-31"
output: html_document
bibliography: references.bib
---

@@

-   Explanation of gender homophily in scholarship

@@

@@

>>>>

---
title: "Draft-Theory"
author: "Verooo"
date: "2024-10-31"
output: html_document
bibliography: references.bib
csl: apa.csl
---

@@

## gender homophily in scholars' collaboration

@@

## h-index and collaboration

@@

# Written references (remove in final paper)

Bornmann, Lutz, Christian Ganser, and Alexander Tekles. 2021. “Simulation of the h Index Use at University Departments Within the Bibliometrics-Based Heuristics Framework: Can the Indicator Be Used to Compare Individual Researchers?” Journal of Informetrics 16 (1): 101237. <https://doi.org/10.1016/j.joi.2021.101237>.

@Bornmann2021

Bornmann, Lutz, Christian Ganser, Alexander Tekles, and Loet Leydesdorff. 2020. “Does the h αα-Index Reinforce the Matthew Effect in Science? The Introduction of Agent-Based Simulations into Scientometrics.” Quantitative Science Studies 1 (1): 331–46.

@Bornmann2020
